---
title: "Chapter 1"
output:
  html_document:
    css: style.css
    highlight: tango
---


```{r, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center", message = "#>")
```

```{r, include=FALSE}
require(data.table)
require(ggplot2)
theme_set(
    theme_bw(base_size = 14, base_family = "Lato") +
    theme(panel.grid = element_line(size = 0.3, color = "gray90"),
          panel.border = element_blank())
)
```

# 1.1 Reading data from a csv file

The workhorse function for reading data from a CSV file is `fread()`. This
function works similarly to `base::read.csv()` or `readr::read_csv()` or
`pandas.read_csv()`, it assumes (one of) the default separator between columns
is `comma (,)`.

We're going to be looking some cyclist data from Montréal. Here's the [original page](http://donnees.ville.montreal.qc.ca/dataset/velos-comptage)
(in French) but it's already included in this repository (`data/bikes.csv` ). We're
using the data from 2012.

This dataset is a list of how many people were on 7 different bike paths in
Montreal, each day.

```{r}
broken_dt <- fread("../data/bikes.csv")
```

Oops, `fread` tell us that it stopped very early on line 17. Other clues include
the separator is semi-colon `;` and there exists missing values. Let take a look
at first few rows and see what we got:

```{r}
head(broken_dt, 3)
```

You'll notice that something wrong with the column names! Fortunately, `fread`
has a bunch of options that will let us fix that, though. Here we'll

* set the column separator explicitly to a `;`

* set the encoding to 'latin1' (the default is 'utf8')

Do it again:

```{r}
bikes <- fread("../data/bikes.csv", sep = ";", encoding = "Latin-1", fill = TRUE)
head(bikes, 3) # looks OK
```

# 1.2 `data.table` object

`data.table`, in its [title](https://github.com/Rdatatable/data.table/blob/master/DESCRIPTION),
call itself an extension of `data.frame` which means two things:

* it inherits all features of `data.frame`, everything you can do with
  `data.frame` is applicable to `data.table`.

* it has its own set of functions to manipulate (aggregate, add, modify, delete,
  join, ...) data efficiently.

```{r}
class(bikes)
is.data.table(bikes)
str(bikes)
```

# 1.3 Selecting a column

There are several ways to select a column from a `data.table` object.

* Like a normal `data.frame`

```{r}
bikes[, "Berri 1"]  # character sub-setting
bikes[, 2L]         # integer sub-setting, not recommended
```
* The `data.table` way

```{r}
bikes[, .(`Berri 1`)] # use backticks to select non-syntactic names
## or
bikes[, list(`Berri 1`)]
## be noted that `data.table` returns vector when you do not use .() or list()
bikes[, `Berri 1`][1:10] # print out the first 10 values
```

# 1.4 Plotting

Unlike Python's `pandas`, `data.table` does not support built-in plotting
feature because it is unnecessary. The R world is rich of packages for making
beautiful graphics.

My favorite tool is `ggplot2`.

```{r}
bikes[, Date := as.Date(Date, "%d/%m/%Y")] # don't worry if you don't understand this line
ggplot(bikes, aes(Date, `Berri 1`)) +
    geom_line(color = "orange")
```
