---
title: "Chapter 1"
output:
  html_document:
    css: style.css
    highlight: tango
---


```{r, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center")
```

```{r, include=FALSE}
require(data.table)
require(ggplot2)
theme_set(
    theme_bw(base_size = 17, base_family = "Lato") +
    theme(panel.grid = element_blank(), panel.border = element_blank())
)
```

# 1.1 Reading data from a csv file

The workhorse function for reading data from a CSV file is `fread`. Similar to
`base::read.csv` or `readr::read_csv` or `pandas.read_csv`, it assumes the
separator between columns is `comma (,)`.

Our example data in this tutorial is cyclist data (source [here](http://donnees.ville.montreal.qc.ca/dataset/velos-comptage) - in
French). Fortunately, Julia has pulled the data which

```{r}
broken_dt <- fread("../data/bikes.csv")
```

```{r}
bikes <- fread("../data/bikes.csv", sep = ";", encoding = "Latin-1", fill = TRUE, index = "Date")
bikes[, Date := as.Date(Date, "%d/%m/%Y")]
head(bikes)
bikes[17]
```


# 1.2 Inspecting the structure of `data.table` object

```{r}
class(bikes)
str(bikes)
```

# 1.3 Selecting a column

There are several ways to select column from a `data.table` object.

* Like a normal `data.frame`

```{r}
bikes[, "Berri 1"] # character sub-setting
bikes[, 2L]         # integer sub-setting, not recommended
```
* The `data.table` way

```{r}
bikes[, .(`Berri 1`)] # same as bikes[, list(`Berri 1`)]
```

Note that `data.table` enables selecting column using bare variable thanks to
meta-programming in R.


# 1.4 Other little known gems of `fread`
