---
title: "Chapter 7"
output:
  html_document:
    css: style.css
    highlight: tango
---


```{r, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center", comment = "#>", message = FALSE)
```

```{r}
require(data.table)
require(ggplot2)
theme_set(
    theme_bw(base_size = 14, base_family = "Lato") +
    theme(panel.grid = element_blank(), panel.border = element_blank())
)
```

One of the main problems with messy data is: how do you know if it's messy or
not?

We're going to use the NYC 311 service request dataset again here, since it's
big and a bit unwieldy.

```{r}
requests <- fread("../data/311-service-requests.csv")
```

# 7.1 How do we know if it's messy?

We're going to look at a few columns here. I know already that there are some
problems with the zip code, so let's look at that first.

To get a sense for whether a column has problems, I usually use `unique()` to
look at all its values. If it's a numeric column, I'll instead plot a histogram
to get a sense of the distribution.

When we look at the unique values in "Incident Zip", it quickly becomes clear
that this is a mess.

Some of the problems:

* Some have been parsed as strings, and some as floats

* There are missing values

* Some of the zip codes are 29616-0759 or 83

* There are some missing values that `data.table` didn't recognize, like 'N/A'
  and 'NO CLUE'

What we can do:

* Normalize 'N/A' and 'NO CLUE' into regular `NA`.

* Look at what's up with the 83, and decide what to do

* Make everything strings

```{r}
requests[, unique(`Incident Zip`)]
```

# 7.2 Fixing the nan values and string/float confusion

We can pass a `na_values` option to `data.table::fread()` to clean this up a
little bit. We can also specify that the type of Incident Zip is a string, not a
float.


```{r}
na_values <- c("NO CLUE", "N/A", "", "NA")
requests <- fread("../data/311-service-requests.csv", na.strings = c(na_values))
requests[, unique(`Incident Zip`)]
```

# 7.3 What's up with the dashes?

```{r}
row_with_dashes <- requests[grepl("-", `Incident Zip`), ]
print(NROW(row_with_dashes))
print(row_with_dashes[, c(1:10)])
```

I thought these were missing data and originally deleted them like this:


But then my friend Dave pointed out that 9-digit zip codes are normal. Let's
look at all the zip codes with more than 5 digits, make sure they're okay, and
then truncate them.

```{r}
long_zip_codes <- requests[nchar(`Incident Zip`) > 5]
long_zip_codes[, unique(`Incident Zip`)]
```

Those all look okay to truncate to me.

```{r}
requests[, `Incident Zip` := substr(`Incident Zip`, 1, 5)]
```

Done

Earlier I thought 00083 was a broken zip code, but turns out Central Park's zip
code 00083! Shows what I know. I'm still concerned about the 00000 zip codes,
though: let's look at that.

```{r}
requests[`Incident Zip` == "00000"]
```

This looks bad to me. Let's set these to `NA`.

```{r}
requests[`Incident Zip` == "00000", `Incident Zip` := NA]
```

Great. Let's see where we are now:

```{r}
requests[, sort(unique(`Incident Zip`))]
```

Amazing! This is much cleaner. There's something a bit weird here, though -- I
looked up 77056 on Google maps, and that's in Texas.

Let's take a closer look:

```{r}
zips <- requests[, `Incident Zip`]
is_close <- startsWith(zips, "0") | startsWith(zips, "1")
is_far <- !is_close & !is.na(zips)
zips[is_far == TRUE]
```

```{r}
requests[is_far, .(`Incident Zip`, Descriptor, City)][order(`Incident Zip`)]
```

Okay, there really are requests coming from LA and Houston! Good to
know. Filtering by zip code is probably a bad way to handle this -- we should
really be looking at the city instead.

```{r}
requests[, .N, by = .(CITY = toupper(City))][order(N, decreasing = TRUE)]
```


It looks like these are legitimate complaints, so we'll just leave them alone.

# 7.4 Putting it together

Here's what we ended up doing to clean up our zip codes, all together:


```{r}
na_values <- c("NO CLUE", "N/A", "NA", "")
requests <- fread("../data/311-service-requests.csv", na.strings = na_values)

fix_zip_codes <- function(zips) {
    zips <- substr(zips, 1, 5)
    zips[zips == "00000"] <- NA
    return(zips)
}

requests[, `Incident Zip` := fix_zip_codes(`Incident Zip`)]
requests[, unique(`Incident Zip`)]
```
