---
title: "Chapter 6"
output:
  html_document:
    css: style.css
    highlight: tango
---


```{r, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center", comment = "#>", message = FALSE)
```

```{r}
require(data.table)
require(ggplot2)
require(gridExtra)
theme_set(
    theme_bw(base_size = 14, base_family = "Lato") +
    theme(panel.grid = element_blank(), panel.border = element_blank())
)
```


We're going to go back to our weather data from Chapter 5, here.

```{r}
weather_2012_final <- fread("../data/weather_2012.csv")
weather_2012_final[, `Date/Time` := as.POSIXct(`Date/Time`)]
str(weather_2012_final)
weather_2012_final[1:5]
```

# 6.1 String Operations

You'll see that the 'Weather' column has a text description of the weather that
was going on each hour. We'll assume it's snowing if the text description
contains "Snow".

We are going to use string processing functions from R's `base`.

```{r}
weather_description <- weather_2012_final[, .(`Date/Time`, Weather)]
is_snowing <- weather_description[grepl("Snow", Weather)]
head(is_snowing)

## let us see when there is snow of the year
ggplot(is_snowing, aes(`Date/Time`, 1)) +
    geom_jitter(shape = 21, color = "white", fill = "palegreen4", size = 2) +
    labs(x = NULL, y = NULL) +
    theme(axis.title.y.left = element_blank(),
          axis.text.y.left = element_blank(),
          axis.ticks.y.left = element_blank(),
          panel.grid = element_line(size = 0.3, color = "gray90"))
```

# 6.2 Use resampling to find the snowiest month

```{r}

monthly_temp <- weather_2012_final[, .(median_temp = median(`Temp (C)`)), by = .(month(`Date/Time`))]
print(monthly_temp)

p1 <- ggplot(monthly_temp, aes(factor(month), median_temp)) +
    geom_col(fill = "darkorange") +
    labs(x = NULL, y = "temperature (C)", title = "Temperature") +
    theme(panel.grid.major.y = element_line(size = 0.3, color = "gray90"))
p1
```

Unsurprisingly, July and August are the warmest.

Now we want to know what is the percentage of time it was snowing each month.

```{r}
monthly_snow <- weather_description[, .(snowing_freq = mean(grepl("Snow", Weather))), by = .(month(`Date/Time`))]
print(monthly_snow)

p2 <- ggplot(monthly_snow, aes(factor(month), snowing_freq)) +
    geom_col(fill = "palegreen4") +
    labs(x = NULL, y = NULL, title = "Snowiness") +
    theme(panel.grid.major.y = element_line(size = 0.3, color = "gray90"))
p2
```

So now we know! In 2012, December was the snowiest month. Also, this graph
suggests something that I feel -- it starts snowing pretty abruptly in November,
and then tapers off slowly and takes a long time to stop, with the last snow
usually being in April or May.


# 6.3 Plotting temperature and snowiness stats together

We can also combine these two statistics (temperature, and snowiness) into one
dataframe and plot them together:

```{r}
weather_stats <- merge(monthly_temp, monthly_snow, by = "month")
print(weather_stats)

grid.arrange(p1, p2, ncol = 1)

```
