---
title: "Chapter 8"
output:
  html_document:
    css: style.css
    highlight: tango
---


```{r, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center", comment = "#>", message = FALSE)
```

```{r}
require(data.table)
require(ggplot2)
theme_set(
    theme_bw(base_size = 14, base_family = "Lato") +
    theme(panel.grid = element_blank(), panel.border = element_blank())
)
```

# 8.1 Parsing Unix timestamps

It's not obvious how to deal with Unix timestamps in `data.table` -- it took me
quite a while to figure this out. The file we're using here is a
popularity-contest file I found on my system at `/var/log/popularity-contest`.

Here's an explanation of [how this file works](https://popcon.ubuntu.com/README).

I'm going to hope that nothing in it is sensitive :)

```{r}
popcon <- fread("../data/popularity-contest.txt", skip = 1, sep = " ", header = FALSE, fill = TRUE)
popcon <- popcon[!NROW(popcon)] # remove last row
colnames(popcon) <- c("atime", "ctime", "package_name", "mu_program", "tag")
```

The colums are the access time, created time, package name, recently used
program, and a tag.

```{r}
print(popcon[1:5])
```

```{r}
popcon[, atime := as.POSIXct(as.integer(atime), origin = "1970-01-01")]
popcon[, ctime := as.POSIXct(as.integer(ctime), origin = "1970-01-01")]
```

So now we can look at our `atime` and `ctime` as dates!

```{r}
print(popcon[1:5])
```

Now suppose we want to look at all packages that aren't libraries.

First, I want to get rid of everything with timestamp 0. Notice how we can just
use a string in this comparison, even though it's actually a timestamp on the
inside?

```{r}
popcon <- popcon[atime > as.POSIXct("1970-01-01 00:00:00")]
```

Now we can use base R string abilities to just look at rows where the package
name doesn't contain 'lib'.

```{r}
nonlibraries <- popcon[!grepl("lib", package_name)]
nonlibraries[order(ctime, decreasing = TRUE)][1:10]
```

Okay, cool, it says that I I installed ddd recently. And postgresql! I remember
installing those things. Neat.
