---
title: "Chapter 5"
output:
  html_document:
    css: style.css
    highlight: tango
---


```{r, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center", comment = "#>", message = FALSE)
```

```{r}
require(data.table)
require(ggplot2)
theme_set(
    theme_bw(base_size = 14, base_family = "Lato") +
    theme(panel.grid = element_blank(), panel.border = element_blank())
)
```

# Summary

By the end of this chapter, we're going to have downloaded all of Canada's
weather data for 2012, and saved it to a CSV.

We'll do this by downloading it one month at a time, and then combining all the
months together.

Here's the temperature every hour for 2012!

```{r, fig.width=10}
weather_2012_final <- fread("../data/weather_2012.csv")
str(weather_2012_final)


## `Date/Time` column is in string format. Let convert it into proper datetime format.
weather_2012_final[, `Date/Time` := as.POSIXct(`Date/Time`)]

## and now we visualise one year of temperature data
ggplot(weather_2012_final, aes(`Date/Time`, `Temp (C)`)) +
    geom_line(color = "palegreen4") +
    scale_y_continuous(breaks = seq(-30, 40, 10)) +
    theme(panel.grid = element_line(colour = "gray90", size = 0.3))
```

# 5.1 Downloading one month of weather data

When playing with the cycling data, I wanted temperature and precipitation data
to find out if people like biking when it's raining. So I went to the site for
[Canadian historical weather data](http://climate.weather.gc.ca/index_e.html#access),
and figured out how to get it automatically.

Here we're going to get the data for March 2012, and clean it up

Here's an URL template you can use to get data in Montreal.

```{r}
url_template <- "http://climate.weather.gc.ca/climateData/bulkdata_e.html?format=csv&stationID=5415&Year={year}&Month={month}&timeframe=1&submit=Download+Data"
```

To get the data for March 2013, we need to format it with `month=3, year=2012`.

```{r, eval=FALSE}
url <- glue::glue(url_template, month = 3, year = 2012)
weather_mar2012 <- fread(url, skip = 15, encoding = "Latin-1", header = TRUE)
## warning: the API didn't work anymore
```

# 5.2 Plotting the temperature by hour of a day


# 5.3 Getting the whole year of data


# 5.4 Saving to a CSV
