---
title: "Chapter 4"
output:
  html_document:
    css: style.css
    highlight: tango
---


```{r, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center", comment = "#>", message = FALSE)
```

```{r}
require(data.table)
require(ggplot2)
theme_set(
    theme_bw(base_size = 14, base_family = "Lato") +
    theme(panel.grid = element_blank(), panel.border = element_blank())
)
```

Okay! We're going back to our bike path dataset here. I live in Montreal, and I
was curious about whether we're more of a commuter city or a biking-for-fun city
-- do people bike more on weekends, or on weekdays?


# 4.1 Adding a 'weekday' column to our dataframe

First, we need to load up the data. We've done this before.

```{r}
bikes <- fread("../data/bikes.csv", sep = ";", encoding = "Latin-1", fill = TRUE, index = "Date")
bikes[, Date := as.Date(Date, "%d/%m/%Y")]

## let's look at the dataset again
head(bikes)
```

```{r}
ggplot(bikes, aes(Date, `Berri 1`)) +
    geom_line(color = "palegreen4") +
    theme(panel.grid = element_line(size = 0.3, color = "gray90"))
```


Next up, we're just going to look at the Berri bike path. Berri is a street in
Montreal, with a pretty important bike path. I use it mostly on my way to the
library now, but I used to take it to work sometimes when I worked in Old
Montreal.

So we're going to create a dataframe with just the Berri bikepath in it.

```{r}
berri_bikes <- bikes[, .(Date, `Berri 1`)]
berri_bikes[1:5]
```

Next, we need to add a `weekday` column. There are two ways to perform the
task. Firstly, we can use function `base::weekdays()` which gives us names of
days of the week. Or we can make use of `data.table` built-in utility function
`wday()`. I used both.

```{r}
berri_bikes[, weekday_int := wday(Date)]
berri_bikes[, weekday_str := weekdays(Date)]
print(berri_bikes)
```

# 4.2 Adding up the cyclists by weekday

This turns out to be really easy!

Performing aggregation is one of distinctive feature of `data.table`, it is
similar to SQL `GROUP BY`. You can check out the
[documentation](https://rdatatable.gitlab.io/data.table/articles/datatable-intro.html#a-grouping-using-by)
to learn more.

In this case, `berri_bikes[, .(total_berri = sum('Berri 1')), by = .(weekday_str)]` means
"Group the rows by weekday and then add up all the values with the same weekday".

```{r}
weekday_counts <- berri_bikes[, .(total_berri = sum(`Berri 1`)), by = .(weekday_str)]

## we want an ordered vector of `weekday` which is really helpful for visualising
weekday_counts[, weekday_str := factor(weekday_str,
                                       levels = c("Monday", "Tuesday", "Wednesday", "Thursday",
                                                  "Friday", "Saturday", "Sunday"),
                                       ordered = TRUE)]

ggplot(weekday_counts, aes(weekday_str, total_berri)) +
    geom_col(fill = "goldenrod", width = 0.8) +
    labs(x = NULL, y = NULL)
```

So it looks like Montrealers are commuter cyclists -- they bike much more during
the week. Neat!


# 4.3 Putting it all together

Let's put all that together, to prove how easy it is.

If you want to play around, try changing `sum` to `max`, `median`, or any other
function you like.


```{r, eval=FALSE}
bikes <- fread("../data/bikes.csv", sep = ";", encoding = "Latin-1", fill = TRUE, index = "Date")
bikes[, Date := as.Date(Date, "%d/%m/%Y")]
berri_bikes[, weekday_str := weekdays(Date)]
weekday_counts <- berri_bikes[, .(total_berri = sum(`Berri 1`)), by = .(weekday_str)]
weekday_counts[, weekday_str := factor(weekday_str,
                                       levels = c("Monday", "Tuesday", "Wednesday", "Thursday",
                                                  "Friday", "Saturday", "Sunday"),
                                       ordered = TRUE)]
ggplot(weekday_counts, aes(weekday_str, total_berri)) +
    geom_col(fill = "goldenrod", width = 0.8) +
    labs(x = NULL, y = NULL)
```
