---
title: "Chapter 3"
output:
  html_document:
    css: style.css
    highlight: tango
---


```{r, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center", comment = "#>", message = FALSE)
```

```{r}
require(data.table)
require(ggplot2)
theme_set(
    theme_bw(base_size = 14, base_family = "Lato") +
    theme(panel.grid = element_blank(), panel.border = element_blank())
)
```

Let's continue with our NYC 311 service requests example.

```{r}
complaints <- fread("../data/311-service-requests.csv")
```

# 3.1 Selecting only noise complaints

I'd like to know which borough has the most noise complaints. First, we'll take
a look at the data to see what it looks like:

```{r}
complaints[1:6, 1:6]
```

To get the noise complaints, we need to find the rows where the "Complaint Type"
column is "Noise - Street/Sidewalk". I'll show you how to do that, and then
explain what's going on.

```{r}
noise_complaints <- complaints[`Complaint Type` == "Noise - Street/Sidewalk"]
noise_complaints[1:3, 1:10]
```

If you look at `noise_complaints`, you'll see that this worked, and it only
contains complaints with the right complaint type. But how does this work? Let's
deconstruct it into two pieces.


```{r}
complaints[, `Complaint Type` == "Noise - Street/Sidewalk"][1:10]
```

This is a big array of `TRUE` and `FALSE` , one for each row in our
dataframe. When we index our dataframe with this array, we get just the rows
where our boolean array evaluated to True. It's important to note that for row
filtering by a boolean array the length of our dataframe's index must be the
same length as the boolean array used for filtering.

You can also combine more than one condition with the & operator like this:

```{r}
is_noise <- complaints[, `Complaint Type` == "Noise - Street/Sidewalk"]
in_brooklyn <- complaints[, `Borough` == "BROOKLYN"]
complaints[is_noise & in_brooklyn][1:6, 1:6]

## or if we wanted a few columns
complaints[is_noise & in_brooklyn][1:6, c('Complaint Type', 'Borough', 'Created Date', 'Descriptor')]
```

# 3.2 A digression about numpy arrays


# 3.3 So, which borough has the most noise complaints?

```{r}
noise_complaints <- complaints[`Complaint Type` == "Noise - Street/Sidewalk"]
count_noise_complaints <- noise_complaints[, .(n_noise = .N), by = .(Borough)][order(n_noise, decreasing = TRUE)]

count_noise_complaints <- merge(count_noise_complaints,
                                complaints[, .(n_complaint = .N), by = .(Borough)],
                                by = "Borough")

count_noise_complaints[, p_noise := n_noise / n_complaint]
print(count_noise_complaints)
```

```{r}
ggplot(count_noise_complaints, aes(reorder(Borough, p_noise), p_noise)) +
    geom_col(fill = "palegreen4", width = 0.7) +
    geom_hline(yintercept = seq(0, 0.03, 0.01), color = "white") +
    labs(x = NULL, y = "percent of noise complaints") +
    coord_flip()
```

So Manhattan really does complain more about noise than the other boroughs!
Neat.
