---
title: "Chapter 2"
output:
  html_document:
    css: style.css
    highlight: tango
---

```{r, include=FALSE}
library(knitr)
hook_output <- knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
    lines <- options$output.lines
    if (is.null(lines)) {
        return(hook_output(x, options))  # pass to default hook
    }
    x <- unlist(strsplit(x, "\n"))
    more <- "..."
    if (length(lines)==1) {        # first n lines
        if (length(x) > lines) {
                                        # truncate the output, but add ....
            x <- c(head(x, lines), more)
        }
    } else {
        x <- c(more, x[lines], more)
    }
                                        # paste these lines together
    x <- paste(c(x, ""), collapse = "\n")
    hook_output(x, options)
})
```



```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", comment = "#>", message = FALSE)
```

```{r}
require(data.table)
require(ggplot2)
theme_set(
    theme_bw(base_size = 14, base_family = "Lato") +
    theme(panel.grid = element_blank(), panel.border = element_blank())
)
```

In this chapter we will be learning syntax for performing most common data
manipulation tasks:

* pick variables based on their names

* pick rows based on their values

* aggregate data (counting) by groups

## 2.1 Import data

We're going to use a new dataset here, to demonstrate how to deal with larger
datasets. This is a subset of the of 311 service requests from [NYC Open Data](https://nycopendata.socrata.com/Social-Services/311-Service-Requests-from-2010-to-Present/erm2-nwe9).

```{r}
complaints <- fread("../data/311-service-requests.csv")
sprintf("Shape of the dataset: %d observations and %d variables", NROW(complaints), NCOL(complaints))
```

## 2.2 What's even in it? (the summary)

Let's take a look at the dataset. By default, the `print` function for
`data.table` object only show the first 5 rows and the last 5 rows. (For
illustrative purpose, I already truncated the output)

```{r output.lines = 24}
print(complaints) # long-scrolling
```

## 2.3 Selecting multiple columns and rows

In [chapter 1](http://htmlpreview.github.io/?https://github.com/chuvanan/rdatatable-cookbook/blob/master/cookbook/chapter1-reading-from-a-csv.html),
we've learnt how to select one column. The syntax remains the same for
selecting multiple variables, and the output is always a new `data.table`.

```{r}
## integer indexing (like data.frame)
complaints[, 1:3]
## character indexing (like data.frame)
output1 <- complaints[, c("Agency Name", "Street Name", "Status")]
print(output1)
## bare column names (only works with data.table)
output2 <- complaints[, .(`Agency Name`, `Street Name`, Status)]
identical(output1, output2)
## or more verbose way:
output3 <- complaints[, list(`Agency Name`, `Street Name`, Status)]
identical(output1, output3)
```

To get the first 5 rows of a `data.table`, we can use a slice: `dt[1:5]`

```{r, output.lines = 24}
complaints[1:5] # output is truncated
```

We can combine these to get the first 5 rows of a column:

```{r}
complaints[1:5, .(`Agency Name`)]
```

## 2.4 What's the most common complaint type?


* The base way

```{r}
count_complaint <- tapply(complaints$`Unique Key`, complaints$`Complaint Type`, length)
## or: table(complaints$`Complaint Type`)
print(class(count_complaint))
## noted that the output is an array, you can treat it like a vector (i.e sorting it)
count_complaint <- count_complaint[order(count_complaint, decreasing = TRUE)]
count_complaint[1:10]
```

* The `data.table` way

```{r}
count_complaint2 <- complaints[, .(n = length(`Unique Key`)), by = .(`Complaint Type`)]
count_complaint2 <- count_complaint2[order(n, decreasing = TRUE)]
print(count_complaint2)
## or more concise
count_complain3 <- complaints[, .(n = .N), by = .(`Complaint Type`)][order(n, decreasing = TRUE)]
identical(count_complaint2, count_complain3)
```

```{r}
ggplot(count_complaint2[1:10], aes(reorder(`Complaint Type`, n), n)) +
    geom_col(fill = "red4") +
    labs(x = NULL, y = NULL) +
    coord_flip()
```
