---
title: "Chapter 9"
output:
  html_document:
    css: style.css
    highlight: tango
---


```{r, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, fig.align = "center",
                      comment = "#>", message = FALSE, warning = FALSE)
```

```{r}
require(DBI)
require(data.table)
require(ggplot2)
theme_set(
    theme_bw(base_size = 14, base_family = "Lato") +
    theme(panel.grid = element_blank(), panel.border = element_blank())
)
```

# 9.1 Reading data from SQL databases

So far we've only talked about reading data from CSV files. That's a pretty
common way to store data, but there are many others! HTML, JSON, relational
databases, Excel (!!!), HDF5, Stata, to name a few. In this chapter we'll talk
about reading data from SQL databases.

I am going to introduce you to two new packages `DBI` and `RSQLite` rather than
`data.table` because it was not designed to talk to relational databases, so it
was not able to read or execute SQL.


```{r}
weather_db <- DBI::dbConnect(RSQLite::SQLite(), "../data/weather_2012.sqlite")
dbListTables(weather_db)
res <- dbSendQuery(weather_db, "SELECT * from weather_2012 LIMIT 3")
dbFetch(res)
dbDisconnect(weather_db)
```

# 9.2 Writing to a SQLite database

```{r}
weather_df <- fread("../data/weather_2012.csv")
con <- dbConnect(RSQLite::SQLite(), "../data/test_db.sqlite")
dbExecute(con, "DROP TABLE IF EXISTS weather_2012")
dbWriteTable(con, "weather_2012", weather_df)
dbDisconnect(con)
```

We can now read from the `weather_2012` table in `test_db.sqlite`, and we see
that we get the same data back:

```{r}
con <- dbConnect(RSQLite::SQLite(), "../data/test_db.sqlite")
res <- dbSendQuery(con, "SELECT * from weather_2012 LIMIT 3")
dbFetch(res)
dbDisconnect(con)
```

The nice thing about having your data in a database is that you can do arbitrary
SQL queries. This is cool especially if you're more familiar with SQL. Here's an
example of sorting by the Weather column:

If you have a PostgreSQL database or MySQL database, reading from it works
exactly the same way as reading from a SQLite database.

# 9.3 Connecting to other kinds of database

To connect to a [MySQL](https://db.rstudio.com/databases/my-sql/) database:

```{r, eval=FALSE}
## install.packages("RMariaDB")
library(DBI)
## Connect to my-db as defined in ~/.my.cnf
con <- dbConnect(RMariaDB::MariaDB(), group = "my-db")
```

To connect to a [PostgreSQL](https://db.rstudio.com/databases/postgresql/) database:

```{r, eval=FALSE}
## install.packages("RPostgres")
library(DBI)
## Connect to the default postgres database
con <- dbConnect(RPostgres::Postgres())
```

*Note: For these to work, you will need a working MySQL / PostgreSQL database,
with the correct localhost, database name, etc.*
